---
title: "Aquatic Survey Results"
author: "K. Meyer, K. Bosley"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



```{r}
# load libraries
library(tidyr)
library(dplyr)
library(gt)
library(ggplot2)
library(here)
library(viridis)
library(webshot)

```

## Load QAQC'd data

```{r}
source('DataQAQC.R')
```

# 1.) Summary of responses/what's presented: Include table w/residency status. Total # submissions (include completes and partials), removed non-resident/visitor (include #) surveys for this purpose 

```{r}

Response.type <- c("Complete","Partial - retained","Partial - no data/dropped")

Res.responses <- as.data.frame(table(ResidentSurvey.dat$Response.status))
Nonres.responses <- as.data.frame(table(NonResidentSurvey.dat$Response.status))

Res.responses <- c(Res.responses$Freq[1],Res.responses$Freq[2],length(remove_ids))
Nonres.responses <- c(Nonres.responses$Freq[1],1,3) # manually examined the four non-res partial entries and determined only 1 included information

Total.responses <- cbind(Response.type,Res.responses,Nonres.responses)
colnames(Total.responses) <- c("Survey Status","Resident","Non-resident")
Total.responses <- as.data.frame(Total.responses)

data <- Total.responses
num.rows <- dim(data)[1]

# Create a summary table of total responses
table_gt <- data %>%
  gt() %>%
  # Left-justify the first/row label column data
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_body(columns = colnames(data)[1])
  ) %>%
  # Left-justify the first/row label column name
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_column_labels(columns = colnames(data)[1])
  ) %>%
   # Center the rest of the table data 
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = colnames(data)[2:length(colnames(data))])
  ) %>%
  # Center the rest of the table column names
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_column_labels(columns = colnames(data)[2:length(colnames(data))])
  ) 

# 
# Save the table as an HTML file first
gtsave(table_gt, here::here("tables", "SurveyStatusTable.html"))

# Convert the HTML table to a PNG file
webshot(here::here("tables", "SurveyStatusTable.html"), here::here("tables", "SurveyStatusTable.png"))

# Save as RTF as an alternative, which may be easier to edit in document
gtsave(table_gt, here::here("tables","SurveyStatusTable.rtf"))

```


## Load region and age-specific 2020 census data and create/save summary table

```{r}
# include 2020 census data from Kevin (County GIS)
# summarize census data across all age groups by City.area, and for those over 18; create summary table

Census.dat <- read.csv(here("data","AgeCensusDataDetailed.csv"))

# Define age group columns
all_age_cols <- c("Under_7", "Age_7_12", "Age_13_18", "Age_19_34", "Age_35_54", "Age_55_64", "Age_65_74", "Age_75_plus")
adult_age_cols <- c("Age_19_34", "Age_35_54", "Age_55_64", "Age_65_74", "Age_75_plus")

# Convert to numeric and calculate row-wise totals
Census.dat[all_age_cols] <- lapply(Census.dat[all_age_cols], as.numeric)
Census.dat <- Census.dat %>%
  mutate(
    CensusTotal = rowSums(select(., all_of(all_age_cols)), na.rm = TRUE),
    CensusOver18   = rowSums(select(., all_of(adult_age_cols)), na.rm = TRUE)
  )

# Group and summarize by City.area
CensusCityAreaSummary <- Census.dat %>%
  group_by(City.area) %>%
  summarise(
    CensusTotal = sum(CensusTotal, na.rm = TRUE),
    CensusOver18   = sum(CensusOver18, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    PropCensusTotal = CensusTotal/ sum(CensusTotal),
    PropCensusOver18   = CensusOver18 / sum(CensusOver18)
  ) %>%
  arrange(City.area)

write.csv(CensusCityAreaSummary,file="CensusCityAreaSummary.csv")


#Note from email from Kevin to Diane on 3/11 regarding West End census numbers: "Using the Washington Office of Financial Management (OFM) adjusted redistricting data it brings the West End population down to 458 people or 1.4% with the inmate population moved to last known residence. The unadjusted West End population was 776."

```

## Creates data columns with Geographic, and Geographic and Area-specific weights
```{r}
# Weighting responses by Geography, and Geography and Age combined

# Load census data again (enables running this section independently)
survey_data <- ResidentSurvey.dat
census_data <- read.csv(here("data","AgeCensusDataDetailed.csv"))

# Filter out rows with missing Age or City.area
survey_clean <- survey_data %>%
  filter(!is.na(Age), !is.na(City.area))

# Map survey age bins to census column names
age_bin_map <- c(
  "18-34" = "Age_19_34",
  "35-54" = "Age_35_54",
  "55-64" = "Age_55_64",
  "65-74" = "Age_65_74",
  "75+"   = "Age_75_plus"
)

survey_clean <- survey_clean %>%
  mutate(AgeBin = recode(Age, !!!age_bin_map))

# Count responses by City.area × AgeBin
survey_counts <- survey_clean %>%
  group_by(City.area, AgeBin) %>%
  summarise(SurveyCount = n(), .groups = "drop")

# Reshape census data to long format
census_long <- census_data %>%
  pivot_longer(cols = -City.area, names_to = "AgeBin", values_to = "PopulationCount")

# Join survey counts with census and compute weights
weights_df <- census_long %>%
  left_join(survey_counts, by = c("City.area", "AgeBin")) %>%
  mutate(SurveyCount = ifelse(is.na(SurveyCount) | SurveyCount == 0, 1, SurveyCount),
         Weight = PopulationCount / SurveyCount)

# Join weights back to the survey data
survey_weighted <- survey_clean %>%
  left_join(weights_df %>% select(City.area, AgeBin, Weight), by = c("City.area", "AgeBin"))

```


# 2.) In which city or area do you live?

```{r}
# Question 2 - in which city or area do you live

# Clean City.area field and remove ambiguous entries
Q2.dat <- ResidentSurvey.dat %>%
  filter(City.area != "" & !is.na(City.area)) 

# Define age group columns
all_age_cols <- c("Under7", "Age7_12", "Age13_18", "Age19_34", "Age35_54", "Age55_64", "Age65_74", "Age75_older")
adult_age_cols <- c("Age19_34", "Age35_54", "Age55_64", "Age65_74", "Age75_older")

# Convert to numeric and calculate row-wise totals
Q2.dat[all_age_cols] <- lapply(Q2.dat[all_age_cols], as.numeric)
Q2.dat <- Q2.dat %>%
  mutate(
    TotalRepresented = rowSums(select(., all_of(all_age_cols)), na.rm = TRUE),
    TotalRepOver18   = rowSums(select(., all_of(adult_age_cols)), na.rm = TRUE)
  )

# Group and summarize by City.area
CityAreaSummary <- Q2.dat %>%
  group_by(City.area) %>%
  summarise(
    TotalRepresented = sum(TotalRepresented, na.rm = TRUE),
    TotalRepOver18   = sum(TotalRepOver18, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    PropRepresented = TotalRepresented / sum(TotalRepresented),
    PropRepOver18   = TotalRepOver18 / sum(TotalRepOver18)
  ) %>%
  arrange(City.area)

#write.csv(CityAreaSummary,file="SurveyCityAreaSummary.csv")

#create figures of side-by-side comparisons of census vs survey representation by area
# first plot by ALL age groups
# Prepare the combined dataset
survey_area_allages <- CityAreaSummary %>% 
  select(City.area, PropRepresented) %>% 
  mutate(DataSource = "Survey") %>% 
  rename(Proportion = PropRepresented)

census_area_allages <- CensusCityAreaSummary %>% 
  select(City.area, PropCensusTotal) %>% 
  mutate(DataSource = "Census") %>% 
  rename(Proportion = PropCensusTotal) 

survey_census_area_allages <- rbind(survey_area_allages, census_area_allages)

# Add percentage labels
survey_census_area_allages <- survey_census_area_allages %>%
  mutate(label = paste0(round(Proportion * 100, 1), "%"))

# Create plot with percentage labels
plot <- ggplot(survey_census_area_allages, aes(x = City.area, y = Proportion, fill = DataSource)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) + 
  geom_text(aes(label = label),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 2.5,
            fontface = "bold") +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1), limits = c(0, max(survey_census_area_allages$Proportion) + 0.05)) +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Percent Distribution by Geographic Area",
    subtitle = "Survey Respondents vs. Census",
    x = "",
    y = "Area Distribution (%)",
    fill = "Data Source"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 14, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    legend.text = element_text(size = 12)
  )

# save figure with fixed dimensions to standardize w/others
ggsave(here("figures","Census_Survey_Area_Fig_AllAges.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

# IF NEEDED BY VOTING POPULATION
#create figures of side-by-side comparisons of census vs survey representation by area
# plot by ages over 18 and over
survey_area_over18 <- CityAreaSummary %>% 
  select(City.area,PropRepOver18) %>% 
  mutate(DataSource = "Survey") %>% 
  rename(Proportion = PropRepOver18)

census_area_over18 <- CensusCityAreaSummary %>% 
  select(City.area, PropCensusOver18) %>% 
  mutate(DataSource = "Census")%>% 
  rename(Proportion = PropCensusOver18)

survey_census_area_over18 <- rbind(survey_area_over18,census_area_over18)
  
```

# 3.) - what is your age?

```{r}
# Q3 - what is your age?

# drop ambiguous entries for this question only
Q3.dat <- ResidentSurvey.dat

# calculate individuals by age category
Under_7 <- sum(Q3.dat$Under7,na.rm=TRUE)
Age_7_12 <- sum(Q3.dat$Age7_12,na.rm=TRUE)
Age_13_18 <- sum(Q3.dat$Age13_18,na.rm=TRUE)
Age_19_34 <- sum(Q3.dat$Age19_34,na.rm=TRUE)
Age_35_54 <- sum(Q3.dat$Age35_54,na.rm=TRUE)
Age_55_64 <- sum(Q3.dat$Age55_64,na.rm=TRUE)
Age_65_74 <- sum(Q3.dat$Age65_74,na.rm=TRUE)
Age_75_older <- sum(Q3.dat$Age75_older,na.rm=TRUE)

SurveyAges <- c(Under_7,Age_7_12,Age_13_18,Age_19_34,Age_35_54,Age_55_64,Age_65_74,Age_75_older)

SurveyAgeBins <- c("Under 7","Age 7-12", "Age 13-18","Age 19-34","Age 35-54","Age 55-64","Age 65-74", "75 and over")

CensusAges <- c(sum(Census.dat$Under_7),sum(Census.dat$Age_7_12),sum(Census.dat$Age_13_18),sum(Census.dat$Age_19_34),sum(Census.dat$Age_35_54),sum(Census.dat$Age_55_64),
                   sum(Census.dat$Age_65_74),sum(Census.dat$Age_75_plus))

SurveyAgeDist.df <- data.frame(SurveyAgeBins, SurveyAges, CensusAges)
write.csv(SurveyAgeDist.df,file = "SurveyCensusAgeDistribution.csv")

######################################
# Ensure correct age order (youngest to oldest)
SurveyAgeDist.df$SurveyAgeBins <- factor(
  SurveyAgeDist.df$SurveyAgeBins,
  levels = c("Under 7", "Age 7-12", "Age 13-18", "Age 19-34",
             "Age 35-54", "Age 55-64", "Age 65-74", "75 and over")
)
# Reshape to long format for plotting
SurveyAgeDist_long <- SurveyAgeDist.df %>%
  pivot_longer(cols = c(SurveyAges, CensusAges),
               names_to = "Source",
               values_to = "Count")

# Create the bar plot with uniform fill color and without the legend
plot <- ggplot(SurveyAgeDist.df, aes(x = SurveyAgeBins, y = SurveyAges)) +
  geom_bar(stat = "identity", position = "dodge", fill = viridis(1)) +  # Use viridis color palette
  #geom_text(aes(label = Frequency), vjust = -0.3, size = 5, fontface = "bold") +  # Add text labels on top of bars
  labs(title = "Age Distribution of Surveyed Individuals",
       x = "",
       y = "Number of People") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 12, face = "bold"),
        legend.position = "none",  # Remove the legend
        plot.margin = margin(1, 1, 2, 1))  # Add margin at the bottom to prevent cutting off labels

ggsave(here("figures","AgeDistribution_AllSurveyIndividuals.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

# plot the age distribution alongside the census age distribution
# absolute values 
plot <- ggplot(SurveyAgeDist_long, aes(x = SurveyAgeBins, y = Count, fill = Source)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_text(aes(label = Count),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 4, fontface = "bold") +
  scale_fill_viridis_d(option = "D") +
  labs(title = "Age Distribution of Survey vs Census (Absolute Numbers)",
       x = "",
       y = "Number of People",
       fill = "Data Source") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
        plot.title = element_text(size = 14, face = "bold"),
        axis.title.y = element_text(size = 12, face = "bold"),
        axis.text.y = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 10),
        plot.margin = margin(1, 1, 2, 1, "lines"))

ggsave(here("figures","AgeDistribution_SurveyCensusAbs.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

# plot the age distribution alongside the census age distribution
# RELATIVE values (proportions) 

# Reshape and calculate proportions
SurveyAgeDist_long <- SurveyAgeDist.df %>%
  pivot_longer(cols = c(SurveyAges, CensusAges),
               names_to = "Source",
               values_to = "Count") %>%
  group_by(Source) %>%
  mutate(Proportion = Count / sum(Count))

# Plot proportions
plot <- ggplot(SurveyAgeDist_long, aes(x = SurveyAgeBins, y = Proportion, fill = Source)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  geom_text(aes(label = scales::percent(Proportion, accuracy = 1)),
            position = position_dodge(width = 0.8),
            vjust = -0.3, size = 4, fontface = "bold") +
  scale_fill_viridis_d(option = "D") +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Age Distribution as Proportion of Total",
       x = "",
       y = "Proportion",
       fill = "Data Source") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
        plot.title = element_text(size = 16, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        axis.text.y = element_text(size = 12, face = "bold"),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 11),
        plot.margin = margin(1, 1, 2, 1, "lines"))

ggsave(here("figures","AgeDistribution_SurveyCensusProportions.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

#################################################
```

# creates gt table (formatted)
```{r}
# produces table of results

# Convert the table to a data frame for better handling
data <- as.data.frame(city_area_counts)
colnames(data) <- c("City.area", "Num.responses")

data <- merge(city_area_resp,city_area_total,by="City.area")
colnames(data) <- c("Area","Survey Respondents","Survey Individuals")


# create a table using gt package
# Create a styled table with the required bold and grey lines
table_gt <- data %>%
  gt() %>%
#  cols_label(
#    City.area = "City/Area",
#    Num.responses = "Number of Responses"
#  ) %>%
  # Left-justify the City/Area column data
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_body(columns = "Area")
  ) %>%
  # Left-justify the "City/Area" column name
  tab_style(
    style = cell_text(align = "left"),
    locations = cells_column_labels(columns = "Area")
  ) %>%
  # Make column headers bold
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = c("Area", "Survey Respondents","Survey Individuals"))
  ) %>%
  # Bold the line above the column headers (top of the table)
  tab_style(
    style = cell_borders(sides = "top", weight = 2),
    locations = cells_column_labels(columns = c("Area", "Survey Respondents","Survey Individuals"))
  ) %>%
  # Bold the line below the column headers (under the header row)
  tab_style(
    style = cell_borders(sides = "bottom", weight = 2),
    locations = cells_column_labels(columns = c("Area", "Survey Respondents","Survey Individuals"))
  ) %>%
  # Bold the bottom line of the table (under the last row of data)
  tab_style(
    style = cell_borders(sides = "bottom", weight = 2),
    locations = cells_body(rows = nrow(city_area_df), columns = c("Area", "Survey Respondents","Survey Individuals"))
  ) 

# Save the table as an HTML file first
print(table_gt)
gtsave(table_gt, here::here("tables", "City_Area_Table.html"))

# Convert the HTML table to a PNG file
webshot(here::here("tables", "City_Area_Table.html"), here::here("tables", "City_Area_Table.png"), vwidth = 800, vheight = 600)
```

# 4.) What is your age?

```{r}
# Q4: What is your age? question

Q4.dat <- filter(ResidentSurvey.dat, Age != "")

# Summarize Age Census Data and summarize to align with survey age bins
CensusAgeTotals <- Census.dat %>%
  summarise(
    Under_7 = sum(Under_7, na.rm = TRUE),
    Age_7_12 = sum(Age_7_12, na.rm = TRUE),
    Age_13_18 = sum(Age_13_18, na.rm = TRUE),
    Age_19_34 = sum(Age_19_34, na.rm = TRUE),
    Age_35_54 = sum(Age_35_54, na.rm = TRUE),
    Age_55_64 = sum(Age_55_64, na.rm = TRUE),
    Age_65_74 = sum(Age_65_74, na.rm = TRUE),
    Age_75_plus = sum(Age_75_plus, na.rm = TRUE)
  )


# Count the occurrences of each unique 'City.area'
age_counts <- as.data.frame(table(ResidentSurvey.dat$Age))
colnames(age_counts) <- c("Age", "Respondents")

survey_age_counts <- age_counts %>%
 mutate(Age = as.character(Age),  # Convert to character
        Age = ifelse(Age == "18-34", "18-34*", Age))%>%  # Modify the specific row
        mutate(Proportion = Respondents/sum(Respondents))%>%
        mutate(DataSource = "Survey")

# reformat census age distribution data
census_totals <- c(CensusAgeTotals$Age_19_34,CensusAgeTotals$Age_35_54,CensusAgeTotals$Age_55_64,CensusAgeTotals$Age_65_74,CensusAgeTotals$Age_75_plus)

census_age_counts <- data.frame(survey_age_counts$Age, census_totals)

census_age_counts <- census_age_counts %>%
    rename(Age = survey_age_counts.Age) %>%
    rename(Respondents = census_totals) %>%
    mutate(Proportion = Respondents/sum(Respondents))%>%
    mutate(DataSource = "Census")

survey_census_ages <- rbind(census_age_counts,survey_age_counts)

# Add percentage labels
survey_census_ages <- survey_census_ages %>%
  mutate(label = paste0(round(Proportion * 100, 1), "%"))

#create barplot comparing survey and census age distribution (of respondents)
plot <- ggplot(survey_census_ages, aes(x = Age, y = Proportion, fill = DataSource)) +
  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate the bars
    geom_text(aes(label = label),
            position = position_dodge(width = 0.9),
            vjust = -0.5,
            size = 3.5,
            fontface = "bold") +
  scale_fill_viridis(discrete = TRUE) +  # Colorblind-friendly palette
  labs(title = "Percent Distribution by Age",
       subtitle = "Survey Respondents vs. Census",
       x = "",
       y = "Age Distribution (%)",
       fill = "Data Source") +  # Add a legend title for clarity
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 14, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )
    
# save figure with fixed dimensions to standardize w/others
ggsave(here("figures","Age_Distribution_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)


```
        
#creates gt table (formatted)
```{r}

# create a table using gt package
# Create a styled table with the required bold and grey lines
table_gt <- data %>%
  gt() %>%
  cols_label(
    Age = "Age",
    Respondents = "Responses",
    Census_2020 = "2020 Census",
    SurveyProportion = "Survey Proportion",
    CensusProportion = "Census Proportion"
  ) %>%
  # Left-justify the Age column data
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_body(columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) %>%
  # Left-justify the "Age" column name
  tab_style(
    style = cell_text(align = "center"),
    locations = cells_column_labels(columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) %>%
  # Make column headers bold
  tab_style(
    style = cell_text(weight = "bold"),
    locations = cells_column_labels(columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) %>%
  # Bold the line above the column headers (top of the table)
  tab_style(
    style = cell_borders(sides = "top", weight = 2),
    locations = cells_column_labels(columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) %>%
  # Bold the line below the column headers (under the header row)
  tab_style(
    style = cell_borders(sides = "bottom", weight = 2),
    locations = cells_column_labels(columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) %>%
  # Bold the bottom line of the table (under the last row of data)
  tab_style(
    style = cell_borders(sides = "bottom", weight = 2),
    locations = cells_body(rows = nrow(data), columns = c("Age", "Respondents","Census_2020","SurveyProportion","CensusProportion"))
  ) 

# Save the table as an HTML file first
gtsave(table_gt, here::here("tables", "Age_Distribution_Data.html"))

# Convert the HTML table to a PNG file
webshot(here::here("tables", "Age_Distribution_Data.html"), here::here("tables", "Age_Distribution_Data.png"), vwidth = 800, vheight = 600)



```

# 5.) how important is it to you to have an aquatic facility?
      # - barplot (percentages) w/table (numbers and percentages)
      # - re-rank to be from most to least important; if possible add percentage in text
      # make the text bigger and boldeed
      
```{r}

#remove skipped responses
Q5.dat <- filter(ResidentSurvey.dat,RateNewPoolImportance !="")

Pool.Importance <- table(Q5.dat$RateNewPoolImportance)
Pool.Importance <- as.data.frame(Pool.Importance)
colnames(Pool.Importance) <- c("ImportanceRank","Frequency")

# Reorder by the frequency (from most to least frequent)
# Sort by Frequency in descending order and reorder ImportanceRank
Pool.Importance <- Pool.Importance %>%
  mutate(ImportanceRank = factor(ImportanceRank,levels = c("Very important", "Somewhat important", 
                                            "Not important", "No opinion/more information needed")))%>%
                                              arrange(ImportanceRank)%>%
                                              mutate(Percentage = round(Frequency/sum(Frequency),2)*100)
                                              


# Create the plot
plot <- ggplot(Pool.Importance, aes(x = ImportanceRank, y = Percentage)) +
  geom_bar(stat = "identity", fill = "#49006a") +  
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 5, color = "black", fontface = "bold") +
  scale_y_continuous(limits = c(0, 70), breaks = seq(0, 70, 10)) +
labs(title = "How Important is a New Aquatic Facility?",
       x = "",
       y = "Frequency",
       fill = "") +  # Add a legend title for clarity
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )


ggsave(here("figures","PoolImportance_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

```
      

# 6.) swimming ability question

```{r}
# summarize the data
library(reshape2)
Q6.dat <- ResidentSurvey.dat[,22:53]
Q6.dat <- colSums(Q6.dat,na.rm=TRUE)

# Reshape the data from wide to long format
AgeAbility <- names(Q6.dat)
long.data <- melt(Q6.dat)

#merge into a new data frame
AgeAbility.dat <- data.frame(AgeAbility,long.data$value)

# need to rename some of the AgeAbility variables so ages can be more easily separated from ability
library(stringr)

AgeAbility.dat <- AgeAbility.dat %>%
  mutate(AgeAbility = AgeAbility %>%
           str_replace_all("7_12", "7.12") %>%
           str_replace_all("13_18", "13.18") %>%
           str_replace_all("19_34", "19.34") %>%
           str_replace_all("35_54", "35.54") %>%
           str_replace_all("55_64", "55.64") %>%
           str_replace_all("65_74", "65.74") %>%
           str_replace_all("75_older", "75.older"))


# Split the variable name into AgeGroup and SwimmingAbility
AgeAbility.df <- AgeAbility.dat %>%
  separate(AgeAbility, into = c("AgeGroup", "SwimmingAbility"), sep = "_") %>%
  rename(Number = long.data.value)

# Calculate proportions within each AgeGroup
AgeAbility.df <- AgeAbility.df %>%
  group_by(AgeGroup) %>%
  mutate(Proportion = Number / sum(Number))

# Reorder AgeGroup to ensure "Under7" comes first, followed by "7.12", then the rest
AgeAbility.df$AgeGroup <- factor(AgeAbility.df$AgeGroup, 
                                 levels = c("Under7", "Age7.12","Age13.18","Age19.34","Age35.54","Age55.64","Age65.74", "Age75.older"))

# Reorder SwimmingAbility to ensure 'Non.swimmer' is first, followed by 'Beginner', 'Intermediate', and 'Advanced'
AgeAbility.df$SwimmingAbility <- factor(AgeAbility.df$SwimmingAbility, 
                                        levels = c("Advanced","Intermediate","Beginner", "Non.swimmer"))

# Create the bar plot with proportions and the reordered x-axis
plot <- ggplot(AgeAbility.df, aes(x = AgeGroup, y = Proportion, fill = SwimmingAbility)) +
  geom_bar(stat = "identity") +  # use stat = "identity" to plot proportions directly
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "",y = "Proportion", title = "Swimming Ability by Age Group") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 16, face = "bold"),
    axis.title.y = element_text(size = 14, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )

ggsave(here("figures","Age_Ability_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)
```

## 7.) have you used an aquatic facility in last 2 years?
      - keep as pie chart
      
```{r}

# filter and format data for pie chart

Q7.dat <- ResidentSurvey.dat %>%
      filter(UsedAquaticsPast2Years != "")
Q7.dat <- table(Q7.dat$UsedAquaticsPast2Years)
Q7.df <- data.frame(Q7.dat)
colnames(Q7.df) <- c("Response","Freq")

# Calculate the percentage for each category
Q7.df <- Q7.df %>%
      mutate(Percentage = round(100 * Q7.df$Freq / sum(Q7.df$Freq), 1))

# Create the pie chart with percentages
plot <- ggplot(Q7.df, aes(x = "", y = Percentage, fill = Response)) +
   geom_bar(stat = "identity", width = 1, show.legend = TRUE, color = "black") +
  coord_polar(theta = "y") +
  scale_fill_viridis(discrete = TRUE) +
  theme_void() +
  theme(
    legend.title = element_blank(),  # Remove the legend title
    plot.title = element_text(size = 16, face = "bold"),  # Increase title size
    legend.text = element_text(size = 14)  # Increase legend label size
  ) +
  labs(title = "Pool Usage in the Past 2 Yrs") +
  geom_text(aes(label = paste(Percentage,"%")), 
            position = position_stack(vjust = .5), 
            color = "darkblue", size = 4, fontface = "bold") 

# Print the pie chart
print(plot)

ggsave(here("figures","Pool_Usage_2yrs.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

```

      
# 8.) which ones did you use? 
      Still To-Do:
          - include list of the 'other' pools
          - shorten pool names for legibility
          - consider clarity of labels, etc.

```{r}
#remove skipped responses
Q8.dat <- filter(ResidentSurvey.dat,WhichAquaticFacilities != "")

#add columns for each separated response, if needed to keep responses associated w/individual survey entries - this will still need some cleaning
  # First, split the 'WhichAquaticFacilities' column based on commas
  #split_data <- ResidentSurvey.dat %>%
  #  # Separate 'WhichAquaticFacilities' into multiple columns
  #  separate(WhichAquaticFacilities, into = paste0("place", 1:10), sep = ",", extra = "merge", fill   = #"right")

#create a single column that contains all pools listed across the entire survey
split_responses <- strsplit(as.character(ResidentSurvey.dat$WhichAquaticFacilities), split = ",")
  
# Flatten the list into a single vector
split_vector <- unlist(split_responses)

#separate Other responses
facility_vector <- split_vector[!grepl("^Other \\(Please specify\\)", split_vector) ]
facility_list <- c("Port Angeles SHORE Aquatic Center",
                    "Sequim Y",
                    "Silverdale Y", 
                    "Port Townsend Y/Mountain View Pool", 
                     "Oak Harbor Community Pool",
                     "North Kitsap Community Pool",
                     "Olympic Aquatic Center - Bremerton", 
                      "Bainbridge Island", 
                      "Private/HOA pool (e.g. Port Ludlow")

# Filter the facility_vector based on whether its elements match the desired values
facilities_clean <- facility_vector[facility_vector %in% facility_list]
                  
other_facilities <- ResidentSurvey.dat[
  grepl("^Other \\(Please specify\\)", ResidentSurvey.dat$WhichAquaticFacilities),
]

Facilities <- table(facilities_clean)
Facilities <- as.data.frame(Facilities)
colnames(Facilities) <- c("AquaticFacility","Frequency")

Facilities <- Facilities %>%
  mutate(AquaticFacility = recode(AquaticFacility,
                                  "Port Angeles SHORE Aquatic Center" = "SHORE Aquatic Center",
                                  "Port Townsend Y/Mountain View Pool" = "Mountain View Pool",
                                  "Private/HOA pool (e.g. Port Ludlow" = "Private/HOA Pool"))


# Reorder by the frequency (from most to least frequently used)
# Sort by Frequency in descending order and reorder ImportanceRank
Facilities <- Facilities %>%
  arrange(desc(Frequency)) %>%
    mutate(AquaticFacility = factor(AquaticFacility, levels = AquaticFacility))


# Create the plot
plot <- ggplot(Facilities, aes(x = AquaticFacility, y = Frequency, fill = "AquaticFacility")) +
  geom_bar(stat = "identity", position = "dodge") +  # Dodge to separate the bars
  scale_fill_viridis(discrete = TRUE) +  # Colorblind-friendly palette
  labs(title = "Aquatic Facility Usage",
       x = "",
       y = "Frequency",
       fill = "")+
  theme_minimal() +
 theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
       plot.margin = margin(10, 10, 10, 20)
    )+
  guides(fill = "none")

# Print the plot
print(plot)
ggsave(here("figures","AquaticFacilityUse_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

```

# 10.) aquatic amenities
      # - produce age-based figures as well, may not use in presentation
      # - select top 5 per age group
      # - work on a way to bin household ages (mean household age) - if we have time??
      
```{r}
# Select the Amenities questions only for easier formatting
# Select only the AquaAmenities columns
AquaAmenities.dat <- ResidentSurvey.dat %>%
  select(starts_with("AquaAmenities_")) %>%
  # Automatically rename columns using gsub to format them
  rename_with(~ gsub("AquaAmenities_", "", .), starts_with("AquaAmenities_")) %>%
  select(-UsageRate)%>%
  # Replace underscores with spaces and capitalize appropriately
  rename_with(~ gsub("_", " ", .), everything()) %>%
  rename_with(~ tools::toTitleCase(.), everything())

# Reshape the data into long format
long_data <- AquaAmenities.dat %>%
  pivot_longer(cols = everything(), 
               names_to = "AquaticAmenity", 
               values_to = "Ranking")
               
# Count the occurrences of each ranking for each amenity
count_data <- long_data %>%
  group_by(AquaticAmenity, Ranking) %>%
  count() %>%
  ungroup()
  
# remove blank entries
count_data <- count_data %>%
  filter(Ranking != "")
  
# Calculate the total count for each amenity
total_counts <- count_data %>%
  group_by(AquaticAmenity) %>%
  summarize(Total = sum(n))

# Join the total counts with the count data
count_data <- count_data %>%
  left_join(total_counts, by = "AquaticAmenity") %>%
  mutate(Percentage = n / Total * 100)

# Order the data by the percentage of "Very likely" votes
ordered_amenities <- count_data %>%
  filter(Ranking == "Not likely") %>%
  arrange(Percentage) %>%
  pull(AquaticAmenity)

# Plot the stacked bar chart
plot <- ggplot(count_data, aes(x = factor(AquaticAmenity, levels = ordered_amenities), 
                       y = Percentage, 
                       fill = Ranking)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +  # Use the Viridis color palette for discrete categories
  labs(title = "Aquatic Amenities by Rank",
       x = "",
       y = "Percentage (%)",
       fill = "Ranking") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )
  
# Print the plot
print(plot)
ggsave(here("figures","AquaticAmenities_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)
```


# create aquatic amenity preference plots for each age bin

```{r}
# PRODUCE AQUATIC AMENITY FIGURE BY AGE BIN

# Select only the AquaAmenities columns
AquaAmenities.dat <- ResidentSurvey.dat %>%
  filter(Age == "35-54")%>%
  select(starts_with("AquaAmenities_")) %>%
  # Automatically rename columns using gsub to format them
  rename_with(~ gsub("AquaAmenities_", "", .), starts_with("AquaAmenities_")) %>%
  select(-UsageRate)%>%
  # Replace underscores with spaces and capitalize appropriately
  rename_with(~ gsub("_", " ", .), everything()) %>%
  rename_with(~ tools::toTitleCase(.), everything())

# Reshape the data into long format
long_data <- AquaAmenities.dat %>%
  pivot_longer(cols = everything(), 
               names_to = "AquaticAmenity", 
               values_to = "Ranking")
               
# Count the occurrences of each ranking for each amenity
count_data <- long_data %>%
  group_by(AquaticAmenity, Ranking) %>%
  count() %>%
  ungroup()
  
# remove blank entries
count_data <- count_data %>%
  filter(Ranking != "")
  
# Calculate the total count for each amenity
total_counts <- count_data %>%
  group_by(AquaticAmenity) %>%
  summarize(Total = sum(n))

# Join the total counts with the count data
count_data <- count_data %>%
  left_join(total_counts, by = "AquaticAmenity") %>%
  mutate(Percentage = n / Total * 100)

# Order the data by the percentage of "Very likely" votes
ordered_amenities <- count_data %>%
  filter(Ranking == "Not likely") %>%
  arrange(Percentage) %>%
  pull(AquaticAmenity)

# Plot the stacked bar chart
plot <- ggplot(count_data, aes(x = factor(AquaticAmenity, levels = ordered_amenities), 
                       y = Percentage, 
                       fill = Ranking)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +  # Use the Viridis color palette for discrete categories
  labs(title = "Aquatic Amenities by Rank, Ages 18-34",
       x = "",
       y = "Percentage (%)",
       fill = "Ranking") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )
  
ggsave(here("figures","AquaAmenities_Age18_34_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)


```


# plot heatmap/matrix to show what percentage of respondents within an age category would be very likely to use each aquatic amenity
```{r}

# Step 1: Filter AquaAmenities, clean names
AquaAmenities.dat <- ResidentSurvey.dat %>%
    select(Age, starts_with("AquaAmenities_")) %>%
    rename_with(~ gsub("AquaAmenities_", "", .), starts_with("AquaAmenities_")) %>%
    select(-UsageRate) %>%
    rename_with(~ gsub("_", " ", .), everything()) %>%
    rename_with(~ tools::toTitleCase(.), everything())

# Step 2: Pivot longer and filter "Very likely" responses
very_likely_data <- AquaAmenities.dat %>%
    pivot_longer(cols = -Age, names_to = "Amenity", values_to = "Ranking") %>%
    filter(Ranking == "Very likely", !is.na(Age), !is.na(Amenity)) %>%
    group_by(Age, Amenity) %>%
    summarise(VeryLikely = n(), .groups = "drop")

# Step 3: Total responses per Age x Amenity combo
total_responses <- AquaAmenities.dat %>%
    pivot_longer(cols = -Age, names_to = "Amenity", values_to = "Ranking") %>%
    filter(!is.na(Ranking), !is.na(Age), !is.na(Amenity)) %>%
    group_by(Age, Amenity) %>%
    summarise(Total = n(), .groups = "drop")

# Step 4: Join and calculate percent
percent_data <- left_join(very_likely_data, total_responses, by = c("Age", "Amenity")) %>%
    mutate(Percentage = VeryLikely / Total * 100)

# Step 5: Bin percentages into 5% bands, cap at 60%
percent_data <- percent_data %>%
    mutate(
        PercentBin = cut(
            Percentage,
            breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60,65,70,75,80),
            labels = c(
                "0–5%", "6–10%", "11–15%", "16–20%", "21–25%", "26–30%",
                "31–35%", "36–40%", "41–45%", "46–50%", "51–55%", "56–60%", "61–65%", "66–70%", "71–75%", "76–80%"
            ),
            right = TRUE
        ),
        # Reverse the factor levels so higher bins map to brighter viridis colors
        PercentBin = factor(PercentBin, levels = rev(levels(PercentBin)))
    )
# Step 6: Order amenities alphabetically (top to bottom)
percent_data <- percent_data %>%
    mutate(Amenity = fct_rev(fct_inorder(sort(unique(Amenity))[match(Amenity, sort(unique(Amenity)))])))

# Step 7: Plot heatmap
plot <- ggplot(percent_data, aes(x = Age, y = Amenity, fill = PercentBin)) +
    geom_tile(color = "white") +
    geom_text(aes(label = paste0(round(Percentage), "%")), size = 3) +
    scale_fill_viridis_d(
        option = "plasma",
        direction = -1,  # optional: reverse so bright = high
        name = "Very Likely to Use"
    ) +
    labs(
        title = "Most Preferred Aquatic Amenities by Age Group",
        x = "Age Group",
        y = "Aquatic Amenity"
    ) +
    theme_minimal(base_size = 12) +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(face = "bold", size = 14),
        legend.title = element_text(face = "bold")
    )
ggsave(here("figures","AquaAmenities_AllAgesMatrix.png"), plot = plot)


```


# 11.) Non aquatic amenities - exactly the same as above
      
```{r}
# Select the Amenities questions only for easier formatting
# Select only the AquaAmenities columns
RecAmenities.dat <- ResidentSurvey.dat %>%
  select(starts_with("RecAmenities_")) %>%
  # Automatically rename columns using gsub to format them
  rename_with(~ gsub("RecAmenities_", "", .), starts_with("RecAmenities_")) %>%
  select(-UsageRate)%>%
  # Replace underscores with spaces and capitalize appropriately
  rename_with(~ gsub("_", " ", .), everything()) %>%
  rename_with(~ tools::toTitleCase(.), everything())

# Reshape the data into long format
long_data <- RecAmenities.dat %>%
  pivot_longer(cols = everything(), 
               names_to = "RecAmenity", 
               values_to = "Ranking")
               
# Count the occurrences of each ranking for each amenity
count_data <- long_data %>%
  group_by(RecAmenity, Ranking) %>%
  count() %>%
  ungroup()
  
# remove blank entries
count_data <- count_data %>%
  filter(Ranking != "")
  
# Calculate the total count for each amenity
total_counts <- count_data %>%
  group_by(RecAmenity) %>%
  summarize(Total = sum(n))

# Join the total counts with the count data
count_data <- count_data %>%
  left_join(total_counts, by = "RecAmenity") %>%
  mutate(Percentage = n / Total * 100)

# Order the data by the percentage of "Very likely" votes
ordered_amenities <- count_data %>%
  filter(Ranking == "Not likely") %>%
  arrange(Percentage) %>%
  pull(RecAmenity)

# Plot the stacked bar chart
plot <- ggplot(count_data, aes(x = factor(RecAmenity, levels = ordered_amenities), 
                       y = Percentage, 
                       fill = Ranking)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +  # Use the Viridis color palette for discrete categories
  labs(title = "Recreation Amenities by Rank",
       x = "",
       y = "Percentage (%)",
       fill = "Ranking") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )
  
# Print the plot
print(plot)
ggsave(here("figures","RecAmenities_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)


```


# recreational amenity ranking barplots for each age bin
```{r}

# PRODUCE REC AMENITY FIGURE BY AGE BIN

# Select only the AquaAmenities columns
RecAmenities.dat <- ResidentSurvey.dat %>%
  filter(Age == "75+")%>%
  select(starts_with("RecAmenities_")) %>%
  # Automatically rename columns using gsub to format them
  rename_with(~ gsub("RecAmenities_", "", .), starts_with("RecAmenities_")) %>%
  select(-UsageRate)%>%
  # Replace underscores with spaces and capitalize appropriately
  rename_with(~ gsub("_", " ", .), everything()) %>%
  rename_with(~ tools::toTitleCase(.), everything())

# Reshape the data into long format
long_data <- RecAmenities.dat %>%
  pivot_longer(cols = everything(), 
               names_to = "RecAmenity", 
               values_to = "Ranking")
               
# Count the occurrences of each ranking for each amenity
count_data <- long_data %>%
  group_by(RecAmenity, Ranking) %>%
  count() %>%
  ungroup()
  
# remove blank entries
count_data <- count_data %>%
  filter(Ranking != "")
  
# Calculate the total count for each amenity
total_counts <- count_data %>%
  group_by(RecAmenity) %>%
  summarize(Total = sum(n))

# Join the total counts with the count data
count_data <- count_data %>%
  left_join(total_counts, by = "RecAmenity") %>%
  mutate(Percentage = n / Total * 100)

# Order the data by the percentage of "Very likely" votes
ordered_amenities <- count_data %>%
  filter(Ranking == "Very likely") %>%
  arrange(desc(Percentage)) %>%
  pull(RecAmenity)

# Plot the stacked bar chart
plot <- ggplot(count_data, aes(x = factor(RecAmenity, levels = ordered_amenities), 
                       y = Percentage, 
                       fill = Ranking)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_viridis(discrete = TRUE) +  # Use the Viridis color palette for discrete categories
  labs(title = "Recreation Amenities by Rank, Ages 75+",
       x = "",
       y = "Percentage (%)",
       fill = "Ranking") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1,size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold"),
    axis.title.y = element_text(size = 15, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
    )
  
ggsave(here("figures","RecAmenities_Age75+_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)


```

#Produce rec amenities figure to compare across all age bins at once

```{r}

# Step 1: Filter non-aquatic RecAmenities, clean names
RecAmenities.dat <- ResidentSurvey.dat %>%
    select(Age, starts_with("RecAmenities_")) %>%
    rename_with(~ gsub("RecAmenities_", "", .), starts_with("RecAmenities_")) %>%
    select(-UsageRate) %>%
    rename_with(~ gsub("_", " ", .), everything()) %>%
    rename_with(~ tools::toTitleCase(.), everything())

# Step 2: Pivot longer and filter "Very likely" responses
very_likely_data <- RecAmenities.dat %>%
    pivot_longer(cols = -Age, names_to = "Amenity", values_to = "Ranking") %>%
    filter(Ranking == "Very likely", !is.na(Age), !is.na(Amenity)) %>%
    group_by(Age, Amenity) %>%
    summarise(VeryLikely = n(), .groups = "drop")

# Step 3: Total responses per Age x Amenity combo
total_responses <- RecAmenities.dat %>%
    pivot_longer(cols = -Age, names_to = "Amenity", values_to = "Ranking") %>%
    filter(!is.na(Ranking), !is.na(Age), !is.na(Amenity)) %>%
    group_by(Age, Amenity) %>%
    summarise(Total = n(), .groups = "drop")

# Step 4: Join and calculate percent
percent_data <- left_join(very_likely_data, total_responses, by = c("Age", "Amenity")) %>%
    mutate(Percentage = VeryLikely / Total * 100)

# Step 5: Bin percentages into 10% bands, cap at 60%
percent_data <- percent_data %>%
    mutate(PercentBin = cut(
        Percentage,
        breaks = c(-Inf, 10, 20, 30, 40, 50, 60),
        labels = c("0–10%", "11–20%", "21–30%", "31–40%", "41–50%", "51–60%"),
        right = TRUE
    ))
# Step 5: Bin percentages into 5% bands, cap at 60%
percent_data <- percent_data %>%
    mutate(
        PercentBin = cut(
            Percentage,
            breaks = c(-Inf, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60),
            labels = c(
                "0–5%", "6–10%", "11–15%", "16–20%", "21–25%", "26–30%",
                "31–35%", "36–40%", "41–45%", "46–50%", "51–55%", "56–60%"
            ),
            right = TRUE
        ),
        # Reverse the factor levels so higher bins map to brighter viridis colors
        PercentBin = factor(PercentBin, levels = rev(levels(PercentBin)))
    )
# Step 6: Order amenities alphabetically (top to bottom)
percent_data <- percent_data %>%
    mutate(Amenity = fct_rev(fct_inorder(sort(unique(Amenity))[match(Amenity, sort(unique(Amenity)))])))

# Step 7: Plot heatmap
plot <- ggplot(percent_data, aes(x = Age, y = Amenity, fill = PercentBin)) +
    geom_tile(color = "white") +
    geom_text(aes(label = paste0(round(Percentage), "%")), size = 3) +
    scale_fill_viridis_d(
        option = "plasma",
        direction = -1,  # optional: reverse so bright = high
        name = "Very Likely to Use"
    ) +
    labs(
        title = "Most Preferred Recreation Amenities by Age Group",
        x = "Age Group",
        y = "Recreation Amenity"
    ) +
    theme_minimal(base_size = 12) +
    theme(
        axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(face = "bold", size = 14),
        legend.title = element_text(face = "bold")
    )
ggsave(here("figures","RecAmenities_AllAgesMatrix.png"), plot = plot)


```


# 12.) How much more would you use an aquatic facility, if it had preferred amenities?
- also includes Q9 and usage for a recreation facility
    
```{r}

# Create a dataframe manually (or reuse `CurrentUsage.dat` if already created)
usage_df <- data.frame(
  Response = c("More than once a week", "Several times a month", "Less than once a month", "Never"),
  Percentage = c(15, 14, 26, 45)
)

# Ensure factor order
usage_df$Response <- factor(usage_df$Response, levels = c("More than once a week", 
                                                           "Several times a month", 
                                                           "Less than once a month", 
                                                           "Never"))

# Plot
plot <- ggplot(usage_df, aes(x = Response, y = Percentage)) +
  geom_bar(stat = "identity", fill = "#49006a") +
  geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 5, color = "black", fontface = "bold") +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  labs(title = "Current Aquatic Facility Usage Rates",
       x = "",
       y = "Percentage (%)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 12, face = "bold"),
        axis.text.y = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 18, face = "bold"),
  plot.margin = margin(t = 10, r = 10, b = 20, l = 50)  # Top, Right, Bottom, Left
  )

ggsave(here("figures","CurrentAquaticUsageRates_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300) 

###################################################################3
# create a figure with current and expected usage w/aquatic amenities
# Create dataframe with manually entered values (replace with your own data if needed)

usage_comparison <- data.frame(
  Response = rep(c("More than once a week", 
                   "Several times a month", 
                   "Less than once a month", 
                   "Never"), 2),
  Percentage = c(15, 14, 26, 45,   # Current
                 28, 32, 18, 22),  # Expected with Aquatic Amenities
  UsageCategory = rep(c("Current Usage", "Expected w/ Aquatic Amenities"), each = 4)
)

# Factor level ordering for clean x-axis
usage_comparison$Response <- factor(usage_comparison$Response,
                                    levels = c("More than once a week", 
                                               "Several times a month", 
                                               "Less than once a month", 
                                               "Never"))

# Plot
plot <- ggplot(usage_comparison, aes(x = Response, y = Percentage, fill = UsageCategory)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_text(aes(label = paste0(Percentage, "%")),
            position = position_dodge(width = 0.9), 
            vjust = -0.5, size = 4.5, fontface = "bold") +
  scale_y_continuous(limits = c(0, 50), breaks = seq(0, 50, 10)) +
  scale_fill_manual(values = c("#49006a", "#ffd92f")) +
  labs(title = "Aquatic Facility Usage",
       x = "",
       y = "Percentage (%)",
       fill = "Usage Category") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 12, face = "bold"),
        axis.text.y = element_text(size = 12, face = "bold"),
        axis.title.y = element_text(size = 14, face = "bold"),
        plot.title = element_text(size = 18, face = "bold"),
        legend.title = element_text(size = 12, face = "bold"),
        legend.text = element_text(size = 11, face = "bold"),
        plot.margin = margin(t = 10, r = 10, b = 20, l = 80)  # Top, Right, Bottom, Left
  )

ggsave(here("figures","UsageRatePercentChange_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300) 
   
```
# Create Sankey plot to show change in aquatic user sentiment before/after preferred amenities

```{r}
# Prepare data for Sankey: Current vs Expected with Aquatic Amenities

library(ggalluvial)

ResidentSurvey.dat <- read.csv("ResidentSurvey.dat.csv")

# Prepare the data
Sankey_Data <- ResidentSurvey.dat %>%
    select(Current = FacilityUsageRate, Expected = AquaAmenities_UsageRate) %>%
    mutate(
        Current = recode(Current, 
                         "Don't use an aquatic facility" = "Never",
                         "Less than once a month" = "Less than\nonce/month",
                         "Several times a month" = "Several\ntimes/month",
                         "More than once a week" = "More than\nonce/week"),
        Expected = recode(Expected, 
                          "Don't use an aquatic facility" = "Never",
                          "Less than once a month" = "Less than\nonce/month",
                          "Several times a month" = "Several\ntimes/month",
                          "More than once a week" = "More than\nonce/week")
    ) %>%
    filter(!is.na(Current), !is.na(Expected), Current != "", Expected != "") %>%
    group_by(Current, Expected) %>%
    summarise(Freq = n(), .groups = "drop") %>%
    mutate(
        # Flip order so "Never" appears at the bottom
        Current = factor(Current, levels = c("More than\nonce/week", 
                                             "Several\ntimes/month", 
                                             "Less than\nonce/month", 
                                             "Never")),
        Expected = factor(Expected, levels = c("More than\nonce/week", 
                                               "Several\ntimes/month", 
                                               "Less than\nonce/month", 
                                               "Never"))
    )

Sankey_Data <- Sankey_Data %>%
  mutate(stratum_fill = Current)  # Only use this for coloring the flows

# Step 2: Plot
sankey_plot <- ggplot(Sankey_Data,
                      aes(axis1 = Current, axis2 = Expected, y = Freq)) +
  # Colored flows with custom fill variable (not directly Current)
  geom_alluvium(aes(fill = stratum_fill, alpha = stratum_fill), width = 1/6) +

  # Gray stratum blocks (outside aes to keep out of legend)
  geom_stratum(width = 1/6, fill = "gray90", color = "black") +

  # Stratum text
  geom_text(stat = "stratum", aes(label = after_stat(stratum)),
            size = 3.5, fontface = "bold") +

  # Axis labels
  scale_x_discrete(
    limits = c("Current Usage", "Expected Usage"),
    expand = c(.05, .05),
    labels = c("Current Usage", "Expected Usage")
  ) +

  # Custom viridis fill and updated legend title
  scale_fill_viridis_d(
    option = "D",
    direction = -1,
    name = "Usage Rate",
    guide = guide_legend(override.aes = list(alpha = 1))
  ) +

  # Alpha transparency mapped to flow color only
  scale_alpha_manual(values = c(
    "More than\nonce/week" = 0.9,
    "Several\ntimes/month" = 0.8,
    "Less than\nonce/month" = 0.7,
    "Never" = 0.5
  ), guide = "none") +  # Don't show alpha in legend

  # Final formatting
  labs(
    title = "Expected Change in Aquatic Facility Usage",
    y = "Number of Respondents"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title = element_text(face = "bold", size = 16, hjust = 0.5),
    legend.position = "bottom",
    axis.title.x = element_blank(),
    axis.text.x = element_text(face = "bold", size = 13, color = "black")
  )

ggsave(here("figures","UsageRateChange_SankeyPlot.png"), plot = sankey_plot)

```

# Create table of tax support numbers, weighted and unweighted

```{r}

# tax support table

# Clean and recode survey data
TaxSupport_clean <- survey_weighted %>%
  filter(City.area != "", TaxSupport != "") %>%
  filter(TaxSupport %in% c("Yes", "No", "I don't know/need more information")) %>%
  mutate(
    TaxSupport = ifelse(TaxSupport == "I don't know/need more information", 
                        "Need more info", 
                        TaxSupport)
  )

# Raw counts
raw_counts <- TaxSupport_clean %>%
  group_by(City.area, TaxSupport) %>%
  summarise(RawCount = n(), .groups = "drop") %>%
  pivot_wider(names_from = TaxSupport, values_from = RawCount, values_fill = 0)

# Weighted counts
weighted_counts <- TaxSupport_clean %>%
  group_by(City.area, TaxSupport) %>%
  summarise(WeightedCount = sum(Weight, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = TaxSupport, values_from = WeightedCount, values_fill = 0)

# Rename weighted columns for clarity
colnames(weighted_counts)[-1] <- paste0(colnames(weighted_counts)[-1], "_Weighted")

# Join raw and weighted tables
summary_table <- left_join(raw_counts, weighted_counts, by = "City.area")

write.csv(summary_table, file="TaxSupportSummaryTable.csv")
```


# characterize individuals wioth most extreme changes in usage patters with/without preferred amenities 

```{r}

# Number of respondents who answere BOTH before and after aquatic amenities usage question
Total <- dim(filter(ResidentSurvey.dat,FacilityUsageRate!="" & AquaAmenities_UsageRate !=""))[1] 

###################################################################################
# examine responses that moved from more than once/week usage to less than once/month usage
LessUsage <- filter(ResidentSurvey.dat, FacilityUsageRate == "More than once a week" & (AquaAmenities_UsageRate == "Never"|AquaAmenities_UsageRate =="Less than once a month"))

LU_ID <- LessUsage$Response.ID

LessUsage_Comments <- filter(ResidentComment.dat, Response.ID %in% LU_ID)
LessUsage_Data <- data.frame(LessUsage$Response.ID,LessUsage$City.area,LessUsage$Age,LessUsage_Comments$Comments)
colnames(LessUsage_Data) <- c("Response.ID","City.area","Age","Comments")

Num.Decreased <- dim(LessUsage_Data)[1]
PercentDecrease <- Num.Decreased/Total

# calculate percentage of those who responded who decreased usage (as defined here)
write.csv(LessUsage_Data,file="LessAquaticUsage_AfterAmenities.csv")

#################################################################################
# examine responses that moved from NEVER usage to the two most frequent categories

MoreUsage <- filter(ResidentSurvey.dat, FacilityUsageRate == "Don't use an aquatic facility" & (AquaAmenities_UsageRate == "More than once a week"|AquaAmenities_UsageRate =="Several times a month"))

MoreUsage_ID <- MoreUsage$Response.ID

MoreUsage_Comments <- filter(ResidentComment.dat, Response.ID %in% MoreUsage_ID)

MoreUsage_Data <- data.frame(MoreUsage$Response.ID,MoreUsage$City.area,MoreUsage$Age,MoreUsage_Comments$Comments)
colnames(MoreUsage_Data) <- c("Response.ID","City.area","Age","Comments")

Num.Increased <- dim(MoreUsage_Data)[1]
PercentIncrease <- Num.Increased/Total

write.csv(MoreUsage_Data,file="MoreAquaticUsage_AfterAmenities.csv")

```


# 13.) non-aquatic amenities

#bar chart of non-aquatic usage rates

```{r}
# Filter and summarize non-aquatic usage data
NonAquaticUsage.dat <- ResidentSurvey.dat %>%
    filter(RecAmenities_UsageRate %in% c("More than once a week", "Several times a month", "Less than once a month", "Never")) %>%
    count(RecAmenities_UsageRate) %>%
    mutate(
        Percentage = round(n / sum(n) * 100, 1),
        Response = factor(RecAmenities_UsageRate, levels = c("More than once a week", 
                                                          "Several times a month", 
                                                          "Less than once a month", 
                                                          "Never"))
    ) %>%
    select(Response, Count = n, Percentage)

# View summary table
print(NonAquaticUsage.dat)

# Plot
plot <- ggplot(NonAquaticUsage.dat, aes(x = Response, y = Percentage)) +
    geom_bar(stat = "identity", fill = "#49006a") +
    geom_text(aes(label = paste0(Percentage, "%")), vjust = -0.5, size = 5, color = "black", fontface = "bold") +
    scale_y_continuous(limits = c(0, max(NonAquaticUsage.dat$Percentage) + 10), breaks = seq(0, 100, 10)) +
    labs(title = "Expected Usage Rates for Non-Aquatic Amenities",
         x = "",
         y = "Percentage (%)") +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 12, face = "bold"),
          axis.text.y = element_text(size = 12, face = "bold"),
          axis.title.y = element_text(size = 14, face = "bold"),
          plot.title = element_text(size = 18, face = "bold"),
          plot.margin = margin(t = 10, r = 10, b = 20, l = 50))

# Save figure
ggsave(here("figures", "CurrentNonAquaticUsageRates_Fig.png"),
       plot = plot, width = 8, height = 6, units = "in", dpi = 300)
```


  
# 14.) tax Q

# create pie charts of tax support, with and without weighting

```{r}
tax.dat <- ResidentSurvey.dat %>%
            select(Response.ID,Age,City.area,RateNewPoolImportance,TaxSupport)

#Create figure of raw response to tax question
cleaned_data <- tax.dat %>%
  filter(TaxSupport %in% c("Yes", "No", "I don't know/need more information")) %>%
  filter(TaxSupport != "" & TaxSupport != "I am not a Jefferson County voter")

# Calculate the percentage of each response in the TaxSupport column
tax_support_counts <- cleaned_data %>%
  count(TaxSupport) %>%
  mutate(percentage = n / sum(n) * 100)

# Filter the data to remove unwanted responses (empty strings and non-voters)
cleaned_data <- tax.dat %>%
  filter(TaxSupport %in% c("Yes", "No", "I don't know/need more information")) %>%
  filter(TaxSupport != "" & TaxSupport != "I am not a Jefferson County voter") %>%
  mutate(TaxSupport = ifelse(TaxSupport=="I don't know/need more information","Need more info",TaxSupport))
  
# Calculate the percentage of each response in the TaxSupport column
tax_support_counts <- cleaned_data %>%
  count(TaxSupport) %>%
  mutate(
    percentage = n / sum(n) * 100,
    label = paste0(TaxSupport, " (", round(percentage, 1), "%)")
  )

# Plot the pie chart with percentages as text and white outlines
plot <- ggplot(tax_support_counts, aes(x = "", y = percentage, fill = label)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = TRUE, color = "black") +
  coord_polar(theta = "y") +
  scale_fill_viridis(discrete = TRUE, name = "Tax Support") +
  labs(
    title = "Tax Support Distribution",
    subtitle = "Raw, un-weighted"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
  )

ggsave(here("figures","TaxSupport_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300) 


##########################################################3
# WEIGHTED VERSION OF THE FIGURE

# Filter, recode, and apply weights from the weighted dataset
tax_weighted <- survey_weighted %>%
  select(Response.ID, Age, City.area, RateNewPoolImportance, TaxSupport, Weight) %>%
  filter(TaxSupport %in% c("Yes", "No", "I don't know/need more information")) %>%
  mutate(TaxSupport = ifelse(
    TaxSupport == "I don't know/need more information", 
    "Need more info", 
    TaxSupport
  ))

# Calculate weighted percentages and prepare legend labels
tax_support_weighted <- tax_weighted %>%
  group_by(TaxSupport) %>%
  summarise(weighted_total = sum(Weight, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    percentage = weighted_total / sum(weighted_total) * 100,
    label = paste0(TaxSupport, " (", round(percentage, 1), "%)")
  )

plot <- ggplot(tax_support_weighted, aes(x = "", y = percentage, fill = label)) +
  geom_bar(stat = "identity", width = 0.5, show.legend = TRUE, color = "black") +
  coord_polar(theta = "y") +
  scale_fill_viridis(discrete = TRUE, name = "Tax Support") +
  labs(
    title = "Tax Support Distribution",
    subtitle = "Weighted by Age and City/area"
  ) +
  theme_void() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 18, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 16, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
  )

# Save the figure
ggsave(here("figures", "TaxSupport_Fig_Weighted_LegendPercents.png"),
       plot = plot, width = 8, height = 6, units = "in", dpi = 300)

```

# create figures of tax support by geographic area, with and without weigthing

```{r}
# filter out 'other' category temporarily; need to go reclassify as possible
TaxArea_clean.dat <- cleaned_data[!grepl("^Other \\(Please specify\\)", cleaned_data$City.area), ]

# drop the blank entries
TaxArea.dat <- TaxArea_clean.dat %>%              
  filter(City.area != "")%>%
  filter(TaxSupport != "") %>%  
  group_by(City.area, TaxSupport) %>%
  summarise(count = n()) %>% 
  group_by(City.area) %>%
  mutate(percentage = count / sum(count) * 100) %>%
  ungroup()

# Order the data by the "yes" votes
ordered_data <- TaxArea.dat %>%
  filter(TaxSupport == "Yes") %>%
  arrange(desc(percentage)) %>%
  pull(City.area) 

# add West End (no yes votes)
ordered_data <- c(ordered_data,"West End")

  plot <- TaxArea.dat %>%
    ggplot(aes(x = factor(City.area, levels = ordered_data),y = percentage, fill = TaxSupport)) +
        geom_bar(stat = "identity") +  # Create the stacked bar chart
        scale_fill_viridis(discrete = TRUE) +  
  labs(title = "Percentage of Tax Support by Area",
       subtitle = "Raw, Un-weighted",
       x = "", 
       y = "Percentage of Responses",
       fill = "Tax Support") +  # Label the legend
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "bold",hjust = 0.5),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    plot.margin = margin(10, 10, 10, 60)
  ) +
 geom_hline(yintercept = 50, linetype = "dashed", color = "grey", size = 1) 

# Print the plot
print(plot)
ggsave(here("figures","TaxSupport_by_Area_percent_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)

###########################################
# produce figure by numbers

# Order the data by the "yes" votes
ordered_data <- TaxArea.dat %>%
  filter(TaxSupport == "Yes") %>%
  arrange(desc(count)) %>%
  pull(City.area) 

# add West End (no yes votes)
ordered_data <- c(ordered_data,"West End")

  plot <- TaxArea.dat %>%
    ggplot(aes(x = factor(City.area, levels = ordered_data),y = count, fill = TaxSupport)) +
        geom_bar(stat = "identity") +  # Create the stacked bar chart
        scale_fill_viridis(discrete = TRUE) +  
        labs(title = "Tax Support by Area in Numbers",
       x = "City Area", 
       y = "Number of Responses",
       fill = "Tax Support") +  # Label the legend
       theme_minimal() +  # Use a minimal theme
       theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Print the plot
print(plot)
ggsave(here("figures","TaxSupport_by_Area_number_Fig.png"), plot = plot, width = 8, height = 6, units = "in", dpi = 300)


################################################################
# WEIGHTED VERSION OF FIGURE

# Remove "Other" and blank city areas, recode PT outside label
TaxArea_clean <- survey_weighted %>%
  filter(!grepl("^Other \\(Please specify", City.area)) %>%
  filter(City.area != "", TaxSupport != "") %>%
  filter(TaxSupport %in% c("Yes", "No", "I don't know/need more information")) %>%
  mutate(
    TaxSupport = ifelse(TaxSupport == "I don't know/need more information", 
                        "Need more info", 
                        TaxSupport)
  )
  

# Summarize weighted counts
TaxArea_weighted <- TaxArea_clean %>%
  group_by(City.area, TaxSupport) %>%
  summarise(weighted_count = sum(Weight, na.rm = TRUE), .groups = "drop") %>%
  group_by(City.area) %>%
  mutate(percentage = weighted_count / sum(weighted_count) * 100) %>%
  ungroup()

# --- PERCENTAGE PLOT ---

# Order by weighted "Yes" percentage
ordered_percent <- TaxArea_weighted %>%
  filter(TaxSupport == "Yes") %>%
  arrange(desc(percentage)) %>%
  pull(City.area)

# Add West End if not already included
if (!"West End" %in% ordered_percent) {
  ordered_percent <- c(ordered_percent, "West End")
}

plot_percent <- TaxArea_weighted %>%
  ggplot(aes(x = factor(City.area, levels = ordered_percent), y = percentage, fill = TaxSupport)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Percentage of Tax Support by Area",
    subtitle = "Weighted by Age and City/Area",
    x = "", 
    y = "Percentage of Responses",
    fill = "Tax Support"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    axis.title.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, face = "bold",hjust = 0.5),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold"),
    plot.margin = margin(10, 10, 10, 60)
  ) +
  geom_hline(yintercept = 50, linetype = "dashed", color = "grey", size = 1)

ggsave(here("figures", "TaxSupport_by_Area_percent_Weighted.png"), 
       plot = plot_percent, width = 8, height = 6, units = "in", dpi = 300)


# --- COUNT (WEIGHTED) PLOT ---

# Order by weighted count of "Yes"
ordered_count <- TaxArea_weighted %>%
  filter(TaxSupport == "Yes") %>%
  arrange(desc(weighted_count)) %>%
  pull(City.area)

# Add West End if not already included
if (!"West End" %in% ordered_count) {
  ordered_count <- c(ordered_count, "West End")
}

plot_count <- TaxArea_weighted %>%
  ggplot(aes(x = factor(City.area, levels = ordered_count), y = weighted_count, fill = TaxSupport)) +
  geom_bar(stat = "identity") +
  scale_fill_viridis(discrete = TRUE) +
  labs(
    title = "Tax Support by Area in Numbers",
    subtitle = "Weighted by Age and City/Area",
    x = "", 
    y = "Weighted Responses",
    fill = "Tax Support"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 12, face = "bold"),
    axis.text.y = element_text(size = 12, face = "bold"),
    plot.title = element_text(size = 20, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 16, hjust = 0.5, face = "bold"),
    legend.text = element_text(size = 12, face = "bold"),
    legend.title = element_text(size = 14, face = "bold")
  )

ggsave(here("figures", "TaxSupport_by_Area_number_Weighted.png"), 
       plot = plot_count, width = 8, height = 6, units = "in", dpi = 300)


```

